---
title: "Getting Started with rddm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with rdmm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
  
*****

## Drift Diffusion Model

The drift diffusion model requires three main parameters:

* v: the drift rate
* a: boundary separation (the distance between the two boundaries)
* t0: non-decision time (in seconds)

Optional parameters include:

* z: starting point (between 0-1, z = 0.5 is the mid point)
* dc: drift criterion or bias in the drift rate (dc > 0 means faster drift to upper boundary than to lower)
* sv: trial-to-trial variability in the drift rate, v ~ N(v, sv))
* sz: trial-to-trial variability in the starting point, z ~ U(z-sz/2, z+sz/2)
* st0: trial-to-trial variability in the non-decision time, t0 ~ U(t0-st0/2, t0+st0/2)

Other parameters may be included to implement collapsing bounds or an urgency signal. These parameters will only take effect if the `bounds` or `urgency` is set -- please see `help(diffusion_model)`:

* a_prime: the degree of boundary collapse
* tc: the time course of boundary collapse
* kappa: the slope of boundary collapse
* uslope: the slope of the urgency signal
* umag: the magnitude of the urgency signal
* udelay: the delay before the urgency signal begins to increase

#### 1. Simulate some data
```{r setup}
library(rddm)

ddm_simulation = sim_ddm(n = 10000,
                         v = 1,
                         a = 2,
                         t0 = 0.3,
                         z = 0.5)

print(head(ddm_simulation$behavior))

```

#### 2. Instantiate a model

Data for the diffusion model must have the following two columns:

* response (0 for lower, 1 for upper boundary)
* rt (in seconds)

This is sufficient if the upper boundary is always the correct answer. If sometimes the lower boundary is the correct answer, please include a third column:

* correctSide (0 for lower, 1 for upper boundary)


```{r ddm_simulate}
ddm_simulation$behavior$correctSide = 1

ddm = diffusion_model$new(ddm_simulation$behavior, # your data
                          model_name = "my_ddm",
                          include = c("z"), # specify optional parameters using the include argument
                          verbose = TRUE) # set to FALSE to ignore the benign parameter messages
```

#### 3. Fit the model to data

For options, please see `help(fit_diffusion_model)`. Many arguments to alter the fitting method are passed on to `modelfitr::fit_model`. Please see `help(fit_model)` for more information on all the package and methods available.
```{r ddm_instantiate}
ddm$fit(method="Nelder-Mead")
```

#### 4. Evaluate model fit
```{r ddm_evaluate}
# Examine parameters and model fit metrics
print(ddm$solution)

# Generate model-predicted behavior
model_predictions = ddm$predict()
print(head(model_predictions))
