---
title: "Getting Started with rddm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with rdmm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*****

The goal of rddm is to provide a user-friendly syntax to work with diffusion models of decision-making.
rddm is centered around two models: the drift diffusion model (DDM; e.g., Ratcliff et al., 1978, Bogacz et al., 2006)
and the pulse diffusion model. 

* Drift Diffusion Model and it's variants:
  + Pure DDM: using only drift rate, boundary, non-decision time, and starting point)
  + Extended DDM: The pure DDM + trial-to-trial variabliity in the drift rate, non-decision time, and starting point
  + Collapsing Bounds: the bounds collapse over time according to a linear, hyperbolic ratio, or weibull function
  + Urgency Signal: a linear or logistic urgency signal
<br>
<br>
* Pulse Diffusion Model: an extension of the DDM that considers moment-by-moment, time-varying evidence, 
such as a train of pulses of sound or light. This model is adapted from Brunton et. al., 2013.
As implemented here, the pulse diffusion model is the exact same weiner diffusion process as the DDM, 
but the drift is a function of the momentary stimulus or evidence towards the upper and lower boundary.
Important notes:
  + This model uses the same parameter convention as the DDM, with three exceptions.
  Trial-to-trial variabliity in the drift rate is replaced with moment-to-moment variability in the drift rate.
  Trial-to-trial variability in the starting point uses a normal distribution, whereas it is uniform in the DDM.
  + This is actually an OU-process, not just an weiner diffusion process. There is an additional parameter, labmda. 
  For lambda = 0, this model is equal to a weiner diffusion process, but for lambda != 0, it is an OU-process.
  + This model offers the same collapsing bounds and urgency paramters as the DDM, but it is not recommended to use 
  these parameters if you are also using allowing lambda to vary.
  
  
*****

## Drift Diffusion Model

The drift diffusion model requires three main parameters:

* v: the drift rate
* a: boundary separation (the distance between the two boundaries)
* t0: non-decision time (in seconds)

Optional parameters include:

* z: starting point (between 0-1, z = 0.5 is the mid point)
* sv: trial-to-trial variability in the drift rate, v ~ N(v, sv))
* sz: trial-to-trial variability in the starting point, z ~ U(z-sz/2, z+sz/2)
* st0: trial-to-trial variability in the non-decision time, t0 ~ U(t0-st0/2, t0+st0/2)

Other parameters may be included to implement collapsing bounds or an urgency signal. These parameters will only take effect if the `bounds` or `urgency` is set (please see `help(diffusion_model)`):

* a_prime: the degree of boundary collapse
* tc: the time course of boundary collapse
* kappa: the slope of boundary collapse
* uslope: the slope of the urgency signal
* umag: the magnitude of the urgency signal
* udelay: the delay before the urgency signal begins to increase

#### 1. Simulate some data
```{r setup}
library(rddm)

ddm_simulation = sim_ddm(n = 10000,
                         v = 1,
                         a = 2,
                         t0 = 0.3,
                         z = 0.5)

print(head(ddm_simulation$behavior))

```

#### 2. Instantiate a model

Data for the diffusion model must have the following two columns:

* response (0 for lower, 1 for upper boundary)
* rt (in seconds)

This is sufficient if the upper boundary is always the correct answer. If sometimes the lower boundary is the correct answer, please include a third column:

* correctSide (0 for lower, 1 for upper boundary)


```{r ddm_simulate}
ddm_simulation$behavior$correctSide = 1

ddm = diffusion_model$new(ddm_simulation$behavior, # your data
                          model_name = "my_ddm",
                          include = c("z"), # specify optional parameters using the include argument
                          verbose = TRUE) # set to FALSE to ignore the benign messages
```

#### 3. Fit the model to data

For options, please see `help(fit_diffusion_model)`. Many arguments to alter the fitting method are passed on to `modelfitr::fit_model`. Please see `help(fit_model)` for more information on all the package and methods available.
```{r ddm_instantiate}
ddm$fit(method="Nelder-Mead")
```

#### 4. Evaluate model fit
```{r ddm_evaluate}
# Examine parameters and model fit metrics
print(ddm$solution)

# Generate model-predicted behavior
model_predictions = ddm$predict()
print(head(model_predictions))
